Approximate State Abstraction

### 2016-01-31 paper/paper.tex
Lemma 2 presentation and proof now fully written up -- fixed it to actually match Lihong

Inspired by the model-irrelevance abstraction, we consider the approximate case. We let $\phi_{model}$ define a type of abstraction that, for fixed $\epsilon_\mathcal{R}$ and $\epsilon_\mathcal{T}$, satisfies:
\begin{multline}
\ep_{model}(s_1) = \ep_{model}(s_2) \rightarrow \\
\forall_a \left|\mathcal{R}(s_1, a) - \mathcal{R}(s_2, a)\right| \leq \epsilon_\mathcal{R}\ \wedge\ \\
\forall_{a,s'} \left|\mathcal{T}(s_1, a,s') - \mathcal{T}(s_2, a,s')\right| \leq \epsilon_\mathcal{T}
\end{multline}

{\bf Lemma 2:} When $S_A$ is created using a function of the $\ep_{model}$ type:
\begin{equation}
\forall_{s \in S_M} : | V^{\pi^*_G}(s) - V^{\pi^*_{A}}(s) | \leq \frac{{\epsilon_{\mathcal{R}}} + \gamma |\mathcal{S}|\epsilon_{\mathcal{T}}}{(1-\gamma)^3}
\end{equation}

{\bf Proof:}
Given the ground \ac{MDP}, $M_G$, consider the \ac{MDP}, $M_G'$ which is identical to $M_G$ except in its transition and rewards.

\begin{equation}
M_G' =  \langle \mathcal{S}_G, \mathcal{A}, \mathcal{R}_G', \mathcal{T}_G', \gamma \rangle
\end{equation}

Where
\begin{align}
\mathcal{R}_G'(s,a) = R_A(\epM(s), a)\\
\mathcal{T}_G'(s,a, s') = R_A(\epM(s), a, \epM(s'))
\end{align}

We know by \cite{li2006towards} that .

Let the maximum upper bound between optimal behavior in the noisy and ground MDP be called $B$.

\begin{align*}
B_{s} &= \max_{s,a}|Q'(s,a)-Q(s,a)|\\
%B_{s} &= \max_{s,a}|R'(s,a) - R(s,a) + \gamma \sum_{s'}\left[T'(s,a,s')\max_{a'}Q'(s',a')-T(s,a,s')\max_{a'}Q(s', a')\right]|\\
&\forall s' \text{ WLOG } T'(s,a,s')\max_{a'}Q'(s',a') \geq T(s,a,s')\max_{a'}Q(s',a')\\
\end{align*}
Applying max is a non-expansion \enote{Probably want a citation and or explanation for this}:
\begin{multline}
=\max_{s,a}(|R'(s,a)-R(s,a)| + \gamma \sum_{s'} \\
\max_{a'} [(T(s,a,s')+T'(s,a,s')- \\ T(s,a,s'))Q'(s', a')-T(s,a,s')Q(s',a')])
\end{multline}
%&\leq \max_{s,a}|R'_s-R_s|+\max_{s,a}\gamma \sum_{s'} \max_{a'} \left[(T(s,a,s')+\epsilon_{\mathcal{T}})Q'(s',a') - T(s,a,s')Q(s',a')\right]\\
%&\leq \epsilon_{\mathcal{R}}+\max_{s,a}\gamma \sum_{s'} \max_{a'} \left[(T(s,a,s')Q'(s',a')+\epsilon_{\mathcal{T}}Q'(s',a') - T(s,a,s')Q(s',a')\right]\\
%&\leq \epsilon_{\mathcal{R}} + \max_{s,a} \gamma \sum_{s'} \max_{a'} \left[(T(s,a,s')(Q'(s',a')-Q(s',a'))+\epsilon_{\mathcal{T}}Q'(s',a') \right]\\
\begin{align*}
&\leq \epsilon_{\mathcal{R}}+\gamma |\mathcal{S}|\epsilon_{\mathcal{T}}Q'_{\text{max}} + \gamma \sum_{s'}\left[ T(s,a,s') B_s\right]\\
&\leq \epsilon_{\mathcal{R}} + \gamma B_{s} + \gamma |\mathcal{S}|\epsilon_{\mathcal{T}}Q'_{max}
\end{align*}

Applying $Q'_{max}=V'_{max} \leq V_{max}=\frac{1}{1-\gamma}$:
\begin{align*}
B_{s} &\leq 
\frac
{{\epsilon_{\mathcal{R}}} + \gamma |\mathcal{S}|\epsilon_{\mathcal{T}}}
{(1-\gamma)^2}
\end{align*}

Therefore:
\begin{align*}
\forall_{s,a} : |Q'(s,a) - Q(s,a)| &\leq \max_{s,a}|Q'(s,a)-Q(s,a)| \\
&\leq  \frac{{\epsilon_{\mathcal{R}}} + \gamma |\mathcal{S}|\epsilon_{\mathcal{T}} }{(1-\gamma)^2}
\end{align*}

Consequently, from the below $\phi_Q^*$ results, we conclude that:
\begin{equation}
\forall_s : |V^{\pi^*_M}(s) - V^{\pi^*_{M'}}(s)| \leq \frac{{\epsilon_{\mathcal{R}}} + \gamma |\mathcal{S}|\epsilon_{\mathcal{T}}}{(1-\gamma)^3}
\end{equation}

@@

We also note that any abstraction that depends only on the reward function, $\mathcal{R}$, can incur unbounded error (or rather, $\textsc{VMax}$).

@@

To demonstrate the efficacy of these approximate abstraction functions, we test out functions in the $\ep{Q^*}$ abstraction family on a variety of domains. We compare the number of states in the abstract MDP to different values of $\epsilon$, as well as the value difference between the optimal policy in the abstract MDP compared to the ground MDP. More specifically, we relate $\epsilon$ to:
\begin{equation}
|V^{\pi_G^*}(s_{init}) - V^{\pi_A^*}(s_{init})|
\end{equation}
For different values of $\epsilon$.

>>>>

Inspired by the model-irrelevance abstraction, we consider the approximate case. We let $\phi_{model}$ define a type of abstraction that, for fixed $\epsilon$, satisfies:
\begin{multline}
\epM(s_1) = \epM(s_2) \rightarrow \\
\forall_a \left| \mcR(s_1, a) - \mcR(s_2, a)\right| \leq \epsilon \wedge\ \\
\forall_{s_A \in \mcS_A} \left|\sum_{g' \in X(s_A)}\mcT(s_1, a, g') - \mcT(s_2, a,g')\right| \leq \epsilon
\end{multline}
where $X(s_A) = \{ g' \in \mcS | \epM(g') = s_A \}$

\enote{Do we want to make it clear that this is ground transitions and rewards?}

\enote{Note that it is possible }

{\bf Lemma 2:} When $S_A$ is created using a function of the $\ep_{model}$ type:
\begin{equation}
\label{eq:lemma2}
\forall_{s \in S_M} : | V^{\pi^*_G}(s) - V^{\pi^*_{A}}(s) | \leq \frac{2\epsilon + 2\gamma(|\mcS-1|\epsilon)}{(1-\gamma)^3}
\end{equation}

{\bf Proof:} Our proof proceeds as follows: we demonstrate that states grouped aggregated under $\epM$ have Q values all within a particular bound. Consequently $\epM$ abstractions are interpretable as an instantiation of of $\epQ$ with a particular $\epsilon$ -- understanding $\epM$ abstractions allows us to conclude our bound by applying Lemma 1.

Let $B$ stand for the maximum Q difference between any pair of ground states in the same abstract state under $\epM$:

\begin{align*}
B = &\max_{s_A \in \mcS_A, s_1, s_2 \in X(s_A), a} |Q(s_1, a) - Q(s_2, a)|\\
=&\max_{s_A \in \mcS_A, s_1, s_2 \in X(s_A), a}      |R(s_1, a) - R(s_2, a) +\\
& \gamma \sum_{g' \in \mcS_G}(\mcT(s_1,a,g')-\mcT(s_2, a, g'))\max_{a'}Q(g', a')|\\
&\leq \epsilon + \gamma \sum_{s_A \in \mcS_A}\sum_{g' \in \mcS_G}\left[(T(s_1, a, g')-T(s_2, a, g')) 		\max_{a'}Q(g', a')	\right]\\
&\text{By the similarity of transitions of grouped states under $\epM$:}\\
& \leq \epsilon + \gamma \sum_{s_A \in \mcS_A} \epsilon Q_{max} \\
&\leq \epsilon + \gamma|\mcS|\epsilon Q_{max}\\
&\text{Applying 0-1 reward normalization and algebra:}\\
& \leq \frac{\epsilon + \gamma(|\mcS| - 1) \epsilon}{1-\gamma}
\end{align*}

Since the Q values of ground states grouped under $\epM$ are strictly less than $B$, we can understand $\epM$ as a type of $\epQ$ with $\epsilon = B$. Applying Lemma 1, this yields Equation \ref{eq:lemma2}.

@@

We also note that any abstraction that depends only on the reward function, $\mathcal{R}$, can incur unbounded error (or rather, $\textsc{VMax}$).`

@@

To demonstrate the efficacy of these approximate abstraction functions, we test functions in the $\ep{Q^*}$ abstraction family on a variety of domains. We compare the number of states in the abstract MDP to different values of $\epsilon$, as well as the value difference between the optimal policy in the abstract MDP compared to the ground MDP. More specifically, we relate $\epsilon$ to:
\begin{equation}
|V^{\pi_G^*}(s_{init}) - V^{\pi_A^*}(s_{init})|
\end{equation}
For different values of $\epsilon$.
